### 2019-11-04 defesa/tex/chapter/definicoes.tex
wip

A rotated ellipse by $\theta$ can be transformed into a axis-parallel centered at the origin by applying two linear transformations. First, the ellipse can be translated to the origin and then rotated by $-\theta$, the reverse transformations are expressed by \autoref{eq:rotated_ellipse}.

\begin{equation}\label{eq:rotated_ellipse}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}=1.
\end{equation}

Also, the coverage region of an ellipse centered at $c\in\R^2$ with shape parameters $(a, b)\in\R^2_{>0}$ and rotation angle $\theta \in [0, \pi]$ is represented by every point $(x, y) \in \R^2$ that satisfies \autoref{eq:rotated_ellipse_co}.

\begin{equation}\label{eq:rotated_ellipse_co}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}\le 1.
\end{equation}

>>>>

A rotated ellipse by $\theta$ can be transformed into a axis-parallel centered at the origin by applying two linear transformations. First, the ellipse can be translated to the origin and then rotated by $-\theta$, the reverse transformations are expressed by \autoref{eq:rotated_ellipse}.
\begin{equation}\label{eq:rotated_ellipse}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}=1.
\end{equation}
Also, the coverage region of an ellipse centered at $c\in\R^2$ with shape parameters $(a, b)\in\R^2_{>0}$ and rotation angle $\theta \in [0, \pi]$ is represented by every point $(x, y) \in \R^2$ that satisfies \autoref{eq:rotated_ellipse_co}.
\begin{equation}\label{eq:rotated_ellipse_co}
\dfrac{((x-c_x)\cos\theta + (y-c_y)\sin\theta)^2}{a^2}+\dfrac{((x-c_x)\sin\theta - (y-c_y)\cos\theta)^2}{b^2}\le 1.
\end{equation}

\section{Polynomials and their roots}

In this work, only univariate polynomials are used, so the definition and results are focused on this case.

A function $p_n: \mathbb{C} \mapsto \mathbb{C}$ is a $n$-degree polynomial if it can be written as:
\begin{equation}\label{eq:poly}
p_n(x) = \sum_{k=0}^{n} a_k x^k.
\end{equation}
The famous Abel-Ruffini Theorem (a proof can be seen in \citeonline{skopenkov2015}) states that for polynomials of degree higher than four, there is no closed formula to determine their roots. However, a numerical approach exists for higher-degree polynomials.

 In \citeonline[p.~195]{horn} a theorem is presented which says that for every univariate polynomial, there exists a companion matrix, such that its eigenvalues are the zeros of that polynomial. At first it seems that we are just trading for the same problem as, in general, to find the eigenvalues of a matrix, one needs to determine the roots of that matrix's characteristic polynomial. Nevertheless, it turns out that companion matrices are very convenient when it comes to finding eigenvalues. They are what is called Hessenberg matrices, and QR algorithms can be used to find their eigenvalues (a $\bigO(n^2)$ algorithm that does that can be found in \citeonline{bini:2007}).
 For example, given a polynomial of degree $4$ written as \autoref{eq:poly} its companion matrix is given by:

 \begin{equation}
 \left[\begin{array}{ccccc}
 0 & 1 & 0 & 0 & 0\\
 0 & 0 & 1 & 0 & 0\\
 0 & 0 & 0 & 1 & 0\\
 0 & 0 & 0 & 0 & 1\\
 -a_0 & -a_1 & -a_2 & -a_3 & -a_4
 \end{array}\right].
 \end{equation}

\section{Real trigonometric polynomial}

The same definition found in \citeonline[p.~150]{powell} for real trigonometric polynomials is given here. They are also referred to as truncated Fourier Series in \citeonline{boyd:2006} and are given by:
\begin{equation}\label{eq:trig_poly}
p_n(\theta) = \sum_{k=0}^{n} a_k\cos(k\theta) + \sum_{k=1}^{n} b_k\sin(k\theta).
\end{equation}
We say that $p_n$ as defined by \autoref{eq:trig_poly} is a $n$-degree real trigonometric polynomial. An important property is stated in \citeonline[p.~150]{powell}, it says that a $n$-degree polynomial can have up to $2n$ distinct roots on the interval $[0, 2\pi)$. It also says that a function written in the format:
$$\begin{array}{cc}\cos^j{\theta}\sin^k{\theta} & \quad j, k \in \mathbb{Z}_+ \end{array},$$
can be transformed into a real trigonometric polynomial of degree $j+k$. Therefore, the following expression also represents a $n$-degree real trigonometric polynomial:
\begin{equation}\label{eq:trig_poly_2}
\sum_{0 \le j+k \le n}c_{j,k}\cos^j{\theta}\sin^k{\theta},
\end{equation}
for some $\{c_{j,k} \in \R: 0 \le j+k \le n\}$.

### 2019-11-04 defesa/tex/chapter/e3p.tex
wip

According to \citeonline[p.~150]{powell}, any function of the form $\{\cos^j{x}\sin^k{x} : j, k \in \mathbb{N}\}$ can be written as a real trigonometric polynomial of degree $j+k$ which can have up to $2(j+k)$ different roots in the interval $[0, 2\pi)$. The reason to bring up this fact is that $\xi$ can be written as $$\sum_{0 \le j+k \le 6} c_{j,k} \cos^{j}(\theta)\sin^{k}(\theta),$$ for some $\{c_{j,k} \in \R: 0 \le j+k \le 6\}$, which implies that it can be converted to a real trigonometric polynomial of degree $6$ which can have up to $12$ roots.

 To show that, just note that it is possible to write $\norm{\varphi(v, \theta)}^2$ and $A(\theta)^2$ in that form, as it can be seen on \autoref{eq:dd} and \autoref{eq:dd2}. It is also possible to see that the term of higher the degree of $\xi$ is the multiplication of the three squared distances, as $\norm{\varphi(v, \theta)}^2$ has degree $2$, the degree of $\xi$ is $6$.
\begin{align}\label{eq:dd}
	\norm{\varphi(v, \theta)}^2 = (v_x\frac{b}{a}\cos\theta + v_y\frac{b}{a}\sin\theta)^2 + (v_y\cos\theta - v_x\sin\theta)^2\\
	\label{eq:dd2} A(\theta)^2=\dfrac{1}{4}\det\left(
	\begin{array}{cc}
		v_x\frac{b}{a}\cos\theta + v_y\frac{b}{a}\sin\theta&v_y\cos\theta - v_x\sin\theta\\
		w_x\frac{b}{a}\cos\theta + w_y\frac{b}{a}\sin\theta&w_y\cos\theta - w_x\sin\theta
	\end{array}\right)^2
\end{align}

@@

\section{Attempts on solving E3PNT}

In this section we give a brief review of some of the approaches we tried on developing a method for E3PNT.

\subsection{Converting $\xi$ into a polynomial}

Using the identity $x = \tan{\frac{\theta}{2}}$, it is possible to convert $\xi(\theta)$ on \autoref{eq:circumscribed_circle_b} into a univariate real polynomial of degree $12$. The famous Abel-Ruffini Theorem (a proof can be seen in \citeonline{skopenkov2015}) states that for polynomials of degree higher than four, there is no closed formula to determine their roots. Therefore, pursuing a numerical method would be a good choice for this problem.

Perhaps the first idea one would have for this problem would be to use root-finding iterative methods, such as Newton's method, to approximate a root $\hat{x}$, divide the polynomial by $(x-\hat{x})$, and apply the method again until the method does not converge. A result published by \citeonline{mc1}, however, says that iterative methods generally are not convergent for polynomials of degree $4$ or more. Also, the process of dividing the polynomial by $(x-\hat{x})$ carries along a lot of error, which could make the method, even if convergent, find spurious roots.

This next approach is very popular in practice as it is present in external libraries like NumPy for Python \cite{python}. In \citeonline[p.~195]{horn} a theorem is presented which says that for every univariate polynomial, there exists a companion matrix, such that its eigenvalues are the zeros of that polynomial. At first it seems that we are just trading for the same problem as, in general, to find the eigenvalues of a matrix, one needs to determine the roots of that matrix's characteristic polynomial. Nevertheless, it turns out that companion matrices are very convenient when it comes to finding eigenvalues. They are what is called Hessenberg matrices, and QR algorithms can be used to find their eigenvalues (a $\bigO(n^2)$ algorithm that does that can be found in \citeonline{bini:2007}).

This method seemed to work fine, but for some instances it did not find roots which were priorly known. This happened probably because of the transformation that was used to convert $\xi$ into a polynomial--later we introduce a different transformation which is more convenient. It can be very numerically unstable as the tangent function grows to fast for angles that are greater than $\frac{\pi}{4}$. Depending on the instance, there could be roots very close to $0$ and others that are too far from it.

\subsection{Using the conic general equation}

@@

\begin{equation}\label{eq:gen_ellipse}
Ax^2+Bxy+Cy^2+Dy+Ex+F=0
\end{equation}

@@

\section{Two methods for E3PNT}

In this section we introduce two methods that successfully find every solution of E3P on the instances we use for experiments. 
The first one uses an approximation of $\xi$, it is slightly more efficient in respect to running time than the second method, however, it the approximation has only asymptotic guarantees and we can assure that it will always work. The second method converts $\xi$ into a complex polynomial and finds its roots by determining the eigenvalues of its companion matrix.

\subsection{An approximation method}

@@

\subsubsection{Chebyshev interpolation}

@@

\subsection{Transforming $\xi$ into a complex polynomial}

This approach is based on a idea published on \citeonline{boyd:2006} which converts a real trigonometric polynomial into a polynomial with a variable $z = e^{i\theta}$. In our case, we replace $\cos{\theta}$ and $\sin{\theta}$ with the following identities:
\begin{align}
\cos{\theta} = \dfrac{e^{i\theta} + e^{-i\theta}}{2}\\
\sin{\theta} = \dfrac{e^{i\theta} - e^{-i\theta}}{2i}.
\end{align}

It is possible to show that with that substitution and changing the variable to $z=e^{it}$, we obtain a function on the form:
\begin{equation}
g(z)=\sum_{k=-6}^6 c_k z^k.
\end{equation}
Then, multiplying it by

>>>>

It turns out that $\xi$ can be written as a real trigonometric polynomial of degree $6$ in the format given by \autoref{eq:trig_poly_2}, which implies that it can have up to $12$ distinct roots.
 To show that, just note that it is possible to write $\norm{\varphi(v, \theta)}^2$ and $A(\theta)^2$ in that form, as it can be seen on \autoref{eq:dd} and \autoref{eq:dd2}. It is also possible to see that the term of higher the degree of $\xi$ is the multiplication of the three squared distances. As $\norm{\varphi(v, \theta)}^2$ has degree $2$, the degree of $\xi$ is $6$.
\begin{align}\label{eq:dd}
	\norm{\varphi(v, \theta)}^2 = (v_x\frac{b}{a}\cos\theta + v_y\frac{b}{a}\sin\theta)^2 + (v_y\cos\theta - v_x\sin\theta)^2\\
	\label{eq:dd2} A(\theta)^2=\dfrac{1}{4}\det\left(
	\begin{array}{cc}
		v_x\frac{b}{a}\cos\theta + v_y\frac{b}{a}\sin\theta&v_y\cos\theta - v_x\sin\theta\\
		w_x\frac{b}{a}\cos\theta + w_y\frac{b}{a}\sin\theta&w_y\cos\theta - w_x\sin\theta
	\end{array}\right)^2
\end{align}

@@

\section{An attempt using the conic general equation}

@@

\begin{equation}\label{eq:gen_ellipse}
Ax^2+Bxy+Cy^2+Dy+Ex+F=0.
\end{equation}
This equation actually represents any conic, for it to be an ellipse the condition $B^2 -4AC < 0$ must be satisfied.

@@

\section{An approximation method}

@@

\subsection{Chebyshev interpolation}

@@

\section{Converting $\xi$ into a polynomial}

As it was pointed out on \autoref{chapter:definitions}, it is possible to find the roots of a univariate polynomial through an algorithm. For that reason, converting $\xi$ somehow into a polynomial can provide a good way to get the solutions of E3PNT.

The first attempt was using the identity $x = \tan{\frac{\theta}{2}}$. This transforms $\xi$ into a polynomial of degree $12$. At first this method seemed to work fine, however, for some instances it did not find roots which were priorly known. This happened probably because of the transformation that was used to convert $\xi$ into a polynomial. It can be very numerically unstable as the tangent function grows to fast for angles that are greater than $\frac{\pi}{4}$. Because of that, depending on the instance, there could be roots very close to $0$ and others that are too far from it.

The second approach is based on a idea published on \citeonline{boyd:2006} which converts a real trigonometric polynomial into a polynomial with a variable $z = e^{i\theta}$. In our case, we replace $\cos{\theta}$ and $\sin{\theta}$ with the following identities:
\begin{align}
\cos{\theta} = \dfrac{e^{i\theta} + e^{-i\theta}}{2}\\
\sin{\theta} = \dfrac{e^{i\theta} - e^{-i\theta}}{2i}.
\end{align}

It is possible to show that with that substitution and changing the variable to $z=e^{it}$, we obtain a function on the form:
\begin{equation}
g(z)=\sum_{k=-6}^6 c_k z^k.
\end{equation}
Then, multiplying $g$ by $z^6 = e^{6i\theta}$, the following polynomial is obtained:

\begin{equation}
h(z) = z^6 g(z) = \sum_{k=0}^{12} c_k z^k.
\end{equation}

Then, every root of $\xi$ is a root of $h$. Note that the converse is not true, there could be roots of $h$ which are not roots of $\xi$, because a root $\hat{z}$ could have $|\hat{z}| \neq 1$ whereas $|e^{it}|$ is $1$.
